---
title: "Analyze MeadoWatch Data"
author: "Janneke Hille Ris Lambers, Aji John, Meera Sethi, Elli Theobald"
date: "Updated 12/17/2020"
output: html_document
---

#Setup for R script and R markdown
1. Load all libraries
2. Specify string behavior
3. Load packages (should we add all packages here?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = FALSE)
library(optimr)
library(tidyverse)
library(leaflet)
library(lubridate)
library(boot)
library(readr) 
```


#Read in clean data, define wildflower community
This chunk of code reads in clean data, defines the focal species, and merges phenology data with plot-level observations for snow disappearance.

1. First, phenological functions (*HRL_phenology_functions.R*), cleaned phenology data (*PhenoSite_Clean.csv*), and snow disappearance data (*MW_SDDall.csv*) are read in. Clean data come from the data wrangling markdown (see *MW_DataWrangling.Rmd*), SDD data come from a microclimate script (*Microclimate.Rmd*), with missing snow disappearance data filled in (see *MW_SDDmodeling.Rmd*).

2. Phenology data are merged with snowmelt data.

3. Next, PhenoSite_Focal is created by subsetting. For reference:
   A. Reflection Lakes species include: c("ANOC", "CAPA", "ERMO", "ERPE", "LIGR", "LUAR", "MIAL", "PEBR", "POBI", "VASI")
   B. Glacier Basin species include: c("ANAR", "ARLA", "ASLE", "CAMI", "ERGR", "LICA", "LUAR", "MEPA", "PEBR", "POBI", "VASI")
   ++Note that LICA was mistaken for ANAR early on, so there is limited data from both species. MEPA and ANAR don't occur in meadows, and MIAL is rare and of all the species least likely to occur on both trails. Recommended species therefore:
c("ANOC", "ARLA", "ASLE", "CAMI", "CAPA", "ERMO", "ERPE", "LIGR", "LUAR", "PEBR", "POBI", "VASI")

4. From this, *nspp* (number of species) and *nyears* (number of years) are defined, which are used in for loops

```{r}
# read in maximum likelihood functions
source("./HRL_phenology_functions.R")

# Read in clean dataset; merge with SDD data
PhenoSite_0 <- read.csv("data/PhenoSite_Clean.csv", header=TRUE)
SDDdat_0 <- read.csv("data/MW_SDDall.csv", header=TRUE)

# Create new SDD - observed when present, predicted when not
SDDfin <- SDDdat_0$SDD
# This substitutes in predicted SDD where sensors failed = NA
SDDfin[is.na(SDDdat_0$SDD)==TRUE] <- 
  round(SDDdat_0$predSDD[is.na(SDDdat_0$SDD)==TRUE],0)
# this removes unwanted columns, including SDD and pred SDD
SDDdat <- subset(SDDdat_0, select=-c(Site_Num, Site_Loc, calibration,
                                     snow_appearance_date, 
                                     snow_disappearance_date,
                                     snow_cover_duration, minimum_soil_temp,
                                     SDD, predSDD))
# This adds SDDfin back to data frame as SDD
SDDdat$SDD <- SDDfin

# Merge SDD and Pheno data
PhenoSite <- merge(PhenoSite_0,SDDdat, 
                   by=c("Year", "Site_Code", "Transect"))

# Define the wildflower community: which species to include
species <- c("ANOC", "ARLA", "ASLE", "CAMI", "CAPA", "ERMO", "ERGR", "ERPE",
             "LIGR", "LUAR", "PEBR", "POBI", "VASI")

# Create PhenoSite_Focal - data set with just species of interest
specieskeep <- PhenoSite$Species %in% species
PhenoSite_Focal <- PhenoSite[specieskeep,]

# Determine how many years of data, how many species, etc
yrs <- unique(PhenoSite_Focal$Year)
nyrs <- length(yrs)
nspp <- length(species)

```


#Fit phenology model for all species data, using second model
This code fits a predictive phenology model to all observations of flowering for a particular species (function *curvefit_species*), for the same flowering community as defined above. Note that this takes a bit of time to run. The code does the following:

1. Subset data for a particular species.

2. For each species, use a for loop to assemble year-plot specific data for the species in question. This requires extracting year-plot specific data adding three no flowering observations before snowmelt (as was done for model fitting for individual year-plot-species data), and creating variables *days* (vector of DOY for all plots included), *phenophase* (vector of 1's and 0's observed), *SDD* (snow disappearance date at each plot for the observation in question).

3. Fit the maximum likelihood model (defined by *curvefit_species*) to the data (see *HRL_phenology_functions.R*).

4. Saves the parameters to *pars_spp* and the AIC values to *aics_sp*. Parameters include per-species slope and intercept (relating peak to SDD), as well as species specific range and max parameters
    
5. After the for loop, *pars_spp* and *aics_sp* are turned into data frames and saved as .csv files in the data folder.

```{r}
# objects in which to append output
pars_spp <- c() #save parameters
aics_sp <-c() #save AIC of overall model 

# Read in per plot model fits, to use for initial parameters
pars_yrpltsp <- read.csv("data/PerPlotCurves.csv", header=TRUE) 

# Now nested for loops to fit curves for all years, species, plots
for(i in 1:nspp){ #loop for each species, to assemble data
  #extract data for that year
  PhenoSite_Species <- PhenoSite_Focal[PhenoSite_Focal$Species==species[i],]
  
  #assemble data - need to add zeros before SDD, define trail year combos
  yrplt <- paste(PhenoSite_Species$Year, PhenoSite_Species$Site_Code, sep="")
  yrplts <- unique(yrplt)
 
  #now define vectors to fill
  days <- c()
  phenophase <- c()
  SDD <- c()
  
  #Now go through each unique year - plot combination per species
  for(j in 1:length(yrplts)){
     PhenoSite_YearPlotSpecies <- PhenoSite_Species[yrplt[]==yrplts[j],]
     tmpdays <- PhenoSite_YearPlotSpecies$DOY #explanatory variable: DOY 
     tmpphenophase <- PhenoSite_YearPlotSpecies$Flower #yes / no flowers
     tmpSDD <- PhenoSite_YearPlotSpecies$SDD
    
     #remove days when no observations were made; NA in phenophase
     tmpdays <- tmpdays[is.na(tmpphenophase)==FALSE]
     tmpSDD <- tmpSDD[is.na(tmpphenophase)==FALSE]
     tmpphenophase <- tmpphenophase[is.na(tmpphenophase)==FALSE]
     
     #Add observations of no flowering 3 weeks prior to snowmelt
     SDDplt <- min(tmpSDD)
     tmpdays <- c(SDDplt-21, SDDplt-14, SDDplt-7,tmpdays)
     tmpphenophase <- c(0,0,0,tmpphenophase)
     tmpSDD <- c(tmpSDD[1:3],tmpSDD)

     #now append to days, phenophase, SDD
     days <- c(days, tmpdays)
     phenophase <- c(phenophase, tmpphenophase)
     SDD <- c(SDD, tmpSDD)
  }

  #specify initial parameters (note, optim can be fussy)
  optimcoef <- c(25,1)
  spdur <- median(pars_yrpltsp$duration[pars_yrpltsp$species==species[i]])
  spmx <- median(pars_yrpltsp$max[pars_yrpltsp$species==species[i]])

  #Fit model: first time using opm to get best results
  param <- c(optimcoef, spdur, spmx) #rep(spdur, times=nyr), rep(spmx, times=nyr))
  modelsnowmod1 <- opm(param, curvefit_species, 
                      method = c("BFGS","hjn", "Nelder-Mead"),
                      control = list(maxit = 50000))
  print(species[i]); print(modelsnowmod1)
  
  #Now refit using Nelder Mead, using initials from previous
  ind <- which(modelsnowmod1$value==min(modelsnowmod1$value))
  param <- c(modelsnowmod1$p1[ind], modelsnowmod1$p2[ind], 
             modelsnowmod1$p3[ind], modelsnowmod1$p4[ind])
  modelsnowmod <- optim(param, curvefit_species, 
                      method = c("Nelder-Mead"),
                      control = list(maxit = 50000))

  #check in case model didn't converge
  if(modelsnowmod$convergence == 1){print(paste(species[i],"-issues",sep=""))}

  #Save parameters to appropriate file
  tmppars <- c(species[i],modelsnowmod$par)
  pars_spp <- rbind(pars_spp,tmppars)
  
  #save AIC
  tmpaic <- c(species[i],round(2*(modelsnowmod$value+length(param)),2))
  aics_sp <- rbind(aics_sp,tmpaic)
}

#make par_spp_peak, pars_spp_range, and pars_spp_max, data frames
dimnames(pars_spp) <- list(c(), c("species","peakint", "peakSDDslope","range","max"))
dimnames(aics_sp) <- list(c(), c("species","AIC"))

#Change to data frame
pars_spp <- data.frame(pars_spp)
aics_sp <- data.frame(aics_sp)

#change storage type 
pars_spp$species <- as.factor(pars_spp$species)
aics_sp$species <- as.factor(aics_sp$species)
aics_sp$AIC <- as.numeric(aics_sp$AIC)

for(i in 2:4){pars_spp[,i] <- as.numeric(pars_spp[,i])}

head(pars_spp)
head(aics_sp)

#Write fitted parameters
write.csv(pars_spp, "data/SpeciesPeakRangeMax.csv", quote=FALSE,
          row.names=FALSE)
write.csv(aics_sp, "data/AIC_speciesmodel.csv", quote=FALSE,
          row.names=FALSE)

```


#Calculate AIC species model vs. separate parameters for all plots
Note - this shows that assuming separate parameters provides a better model fit, despite needing a lot more parameters. So we can't use model fit to justify fitting the all species model. However, looking below you can see this model does a pretty good job predicting peak flowering, which is the main thing we are interested in.
```{r}
aics_yrpltsp <- read.csv("data/AIC_plotmodel.csv", header = TRUE)
aics_yrpltsp2 <- tapply(aics_yrpltsp$AICalt, aics_yrpltsp$species, sum)
aics_yrpltsp3 <- data.frame(names(aics_yrpltsp2),aics_yrpltsp2)
colnames(aics_yrpltsp3) <- c("species","plotmodel")
rownames(aics_yrpltsp3) <- c()
aics_comp <- merge(aics_sp, aics_yrpltsp3)
dimnames(aics_comp)[[2]][2] <- "snowmodel"
dimnames(aics_comp)[[2]][3] <- "plotmodel"

aics_comp
```

# Show plot-level peak fits vs. fitted coefficients
This graph shows the plot level fits of peak flowering vs. SDD found at the site, and the fitted relationship from the model fit across species (black line). Shown additionally is the snowmelt date (blue dashed line) and the line describing a linear model of the plot level peak parameters vs. SDD.
```{r}

#Read in per plot model fits, so individual plot fits can be used
pars_yrpltsp <- read.csv("data/PerPlotCurves.csv", header=TRUE) 

#extract species
#species <- pars_spp_peak$species
yrcols <- c("green","lightblue","orange","purple","pink","red","grey")

##Plot SDD vs. peak per species
for(i in 1:length(species)){
  sp_pars <- pars_yrpltsp[pars_yrpltsp$species==species[i],]
  yrcolind <- sp_pars$year - 2012
  
  #Plot points showing plot SDD and plot level peak fits
  par(mfrow=c(1,1), omi=c(0,0,0,0), mai=c(0.5,0.6,0.5,0.5), 
      tck=-0.02, mgp=c(1.1,0.5,0))
  plot(sp_pars$SDD, sp_pars$peak, pty="s",
       pch=21, bg=yrcols[yrcolind], cex=1.5,
       xlab="Snow Disappearance Date", ylab="Peak Flowering")
  title(species[i])
  legend(x="topleft", legend=yrs, pt.bg=yrcols, pch=21, 
         cex=0.75, pt.cex=1.5)
  
  #add line showing modeled relationship between SDD and peak
  intSDD <- pars_spp$peakint[pars_spp$species==species[i]]
  slSDD <- pars_spp$peakSDDslope[pars_spp$species==species[i]]
  abline(intSDD, slSDD)
  
  #add line showing linear model between SDD and fit
  SDDcoef <- coef(lm(sp_pars$peak~sp_pars$SDD))
  abline(SDDcoef, col="grey", lty=2, lwd=2)
  
  #add line showing snowmelt
  abline(0,1, col="lightblue", lty=3, lwd=2)
  
  #add legend showing line types
  legend(x="bottomright", legend=c("MLE","LM","snow"), lty=c(1,2,3),
         lwd=c(1,1,2), col=c("black","grey","lightblue"), cex=0.75)
  
}

```
