---
title: "Maximum likelihood models: per year/plot/species"
author: "Janneke Hille Ris Lambers, Aji John, Meera Sethi, Elli Theobald"
date: "12/15/2020"
output: html_document
---

#Setup for R script and R markdown
1. Load all libraries
2. Specify string behavior
3. Load packages (should we add all packages here?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = FALSE)
library(tidyverse)
library(leaflet)
library(lubridate)
library(boot)
library(readr) 
```


#Read in clean phenology data
This chunk of code:

1. Reads in the cleaned up data (from data wrangling script), and defines the years, species, etc considered.

2. Reads in SDD data, merges it with PhenoSite data

```{r}
# Read in tidy data
PhenoSite_0 <- read.csv("data/PhenoSite_Clean.csv", header=TRUE)
SDDdat_0 <- read.csv("data/MW_SDDall.csv", header=TRUE)

# Create new SDD - observed when present, predicted when not
SDDfin <- SDDdat_0$SDD
# This substitutes in predicted SDD where sensors failed = NA
SDDfin[is.na(SDDdat_0$SDD)==TRUE] <- 
  round(SDDdat_0$predSDD[is.na(SDDdat_0$SDD)==TRUE],0)
# this removes unwanted columns, including SDD and pred SDD
SDDdat <- subset(SDDdat_0, select=-c(Site_Num, Site_Loc, calibration,
                                     snow_appearance_date, snow_disappearance_date,
                                     snow_cover_duration, minimum_soil_temp, SDD,
                                     notes,predSDD))
# This adds SDDfin back to data frame as SDD
SDDdat$SDD <- SDDfin

# Merge SDD and Pheno data
PhenoSite <- merge(PhenoSite_0,SDDdat, 
                   by=c("Year", "Site_Code", "Transect"))

# Define years of data, data filter, etc
yrs <- unique(PhenoSite$Year)
nyrs <- length(yrs)
species <- unique(PhenoSite$Species)
nspp <- length(species)
out_thresh <- max(PhenoSite$distyes[PhenoSite$Flower==1])

# Print this out as reminder
print("Number of flowering observations, columns")
print(dim(PhenoSite))
print("Number of years, species, outlier threshold")
print(c(nyrs, nspp, out_thresh))

```


#Model fitting: per plot / year / species
This code does the following:

1. Loads phenology functions (see HRL_phenology_functions.R).

2. For each plot / year / species data, fits null model and phenology curve to yes / no observations of flowering phenology. Model fit parameters (peak, range and maximum parameters) are written to object pars_yrpltsp, and saved.

3. For each plot / year / species combo, calculates AIC's. This is written to aics_yrpltsp, and saved.


```{r}
#read in appropriate functions
source("./HRL_phenology_functions.R")

#define object in which to save parameters, AIC values
pars_yrpltsp <- c() 
aics_yrpltsp <- c() 

#For loops to fit curves per year/plot/species, save parameters, plot 
for(i in 1:nyrs){ # First for loop: runs through years

  #extract data for that year
  PhenoSite_Year <- PhenoSite[PhenoSite$Year==yrs[i],]
  
  #pull out unique plots for year in question
  plots <- unique(PhenoSite$Site_Code)
  
  #Nested for loop for each plot
  for(j in 1:length(plots)){ # Second for loop: each plot
    PhenoSite_YearPlot <- PhenoSite_Year[PhenoSite_Year$Site_Code==plots[j],]

    #Identify species in plot
    spinplt <- unique(PhenoSite_YearPlot$Species)
    if(length(spinplt)==0){next} #break if focal spp not in plot
    
    #Nested for loop for each species
    for(k in 1:length(spinplt)){ #Third for loop: each species in the pot

      #Extract data for the species in question      
      PhenoSite_YearPlotSpecies <- 
        PhenoSite_YearPlot[PhenoSite_YearPlot$Species==spinplt[k],]
      
      #define parameters for curvefitting
      days <- PhenoSite_YearPlotSpecies$DOY #explanatory variable: DOY 
      phenophase <- PhenoSite_YearPlotSpecies$Flower #yes / no flowering
	    
	    #add three weeks of zeroes before earliest SDD in those plots
	    SDDplt <- min(PhenoSite_YearPlotSpecies$SDD)
	    days <- c(SDDplt-21, SDDplt-14, SDDplt-7,days)
	    phenophase <- c(0,0,0,phenophase)

      #now fit null model
      model0 <- optimize(nullfit, c(0.000001,0.999999)) #fit null model
	
      #now fit alternative model - curve
      param <- c(mean(days[phenophase[]==1]), -0.001, 0) # initial parameters
      model1 <- optim(param, curvefit_perplot, control = list(maxit = 50000))
      if(model1$convergence==1){
        print(paste(spinplt[k],"no convergence", sep="-"))}
      
      #write the model tests / AIC to data frame
      AICnull <- round(2*(model0$objective+1),1)
      AICalt <- round(2*(model1$value + 3),1)
      pcurve <- signif(pchisq(model1$value-model0$objective,2),3)

      #save AIC values
      tmp_aic_vals <- c(yrs[i], as.character(plots[j]),
                        as.character(spinplt[k]),AICnull,AICalt)
      
      #save parameters to spyrpltpars
      tmp_pars <- c(yrs[i], as.character(plots[j]), SDDplt, 
                    as.character(spinplt[k]), model1$par[1:3])
      
      pars_yrpltsp <- rbind(pars_yrpltsp, tmp_pars)
      aics_yrpltsp <- rbind(aics_yrpltsp, tmp_aic_vals)
      
    }
  }
}
  
# turn spyrpltpars into a data frame
dimnames(pars_yrpltsp) <- list(c(), c("year", "plot", "SDD", "species",
                                     "peak", "duration", "max"))
pars_yrpltsp <- data.frame(pars_yrpltsp)

#change storage type to numeric - all except plot (since a few plots have a, b)
pars_yrpltsp$year <- as.numeric(pars_yrpltsp$year)
pars_yrpltsp$species <- as.factor(pars_yrpltsp$species)
pars_yrpltsp$SDD <- as.numeric(pars_yrpltsp$SDD)
pars_yrpltsp$peak <- as.numeric(pars_yrpltsp$peak)
pars_yrpltsp$duration <- as.numeric(pars_yrpltsp$duration)
pars_yrpltsp$max <- as.numeric(pars_yrpltsp$max)

# turn aic vector to dataframe
dimnames(aics_yrpltsp) <- list(c(), c("year", "plot", "species", "AICnull", "AICalt"))
aics_yrpltsp <- data.frame(aics_yrpltsp)
aics_yrpltsp$year <- as.numeric(aics_yrpltsp$year)
aics_yrpltsp$species <- as.factor(aics_yrpltsp$species)
aics_yrpltsp$AICnull <- as.numeric(aics_yrpltsp$AICnull)
aics_yrpltsp$AICalt <- as.numeric(aics_yrpltsp$AICalt)

#examine pars, aics
head(pars_yrpltsp)
head(aics_yrpltsp)

#Write output to data folder
write.csv(pars_yrpltsp, "data/PerPlotCurves.csv", quote=FALSE,
          row.names=FALSE)

#Write AIC output to data folder
write.csv(aics_yrpltsp, "data/AIC_plotmodel.csv", quote=FALSE,
          row.names=FALSE)

```


#Graph to visualize peak flowering by year  trail
This code takes estimates of peak flowering from year-plot-species specific model fits (in pars_yrpltsp) and plots those estimates by year (year on the x-axis, DOY on the y axis) and by trail.

```{r}
##Plot peak flowering estimates of all species, trails, plots on one graph
par(mfrow=c(1,1),omi=c(0,0,0,0), mai=c(0.5,0.4,0.4,0.2), 
    tck=-0.01, mgp=c(1.25,0.25,0), xpd=TRUE)

#set plotting colors - 16 total to accommodate all species if needed
plotcol <- c("yellowgreen","magenta","orange","purple","yellow","springgreen",
             "pink","purple","navyblue","azure4","yellow4","orchid",
             "turquoise","salmon","maroon","black")

#create plot to add points to
#earlypk <- min(spyrpltpars$peak); latepk <- max(spyrpltpars$peak)
earlypk <- 135; latepk <- 255

plot(2016,175, xlim=c(2012, 2020), ylim=c(earlypk,latepk),type="n",
     xaxp=c(2012,2020,8), yaxt="n", xlab="Year",ylab="Flowering")
text(2011.5,srt=90, c(135, 165, 195, 225, 255),-0.1, 
     labels=c("May", "Jun", "Jul", "Aug", "Sept"))
  
for(trail in 1:2){
  if(trail==1){
    pars_yrpltsp2 <- pars_yrpltsp[substr(pars_yrpltsp$plot,1,2)=="RL",]}
  if(trail==2){
    pars_yrpltsp2 <- pars_yrpltsp[substr(pars_yrpltsp$plot,1,2)=="GB",]}
  
  #how many years? differs per trail
  yrs2 <- unique(pars_yrpltsp2$year)

  #extract data per year, plot
  for(i in 1:length(yrs2)){
    paryear <- pars_yrpltsp2[pars_yrpltsp2$year==yrs2[i],]
 
    #now pull out all data for a species
    for(j in 1:length(species)){
      paryearsp <- paryear[paryear$species==species[j],]
      if(dim(paryearsp)[1]==0){next}
      pks <- paryearsp$peak
      tiny <- 0; if(trail==2){tiny <- 0.333}
      if(trail==1){pltshp <-21}; if(trail==2){pltshp <- 24}
      points((rep(yrs2[i],length(pks))+jitter(rep(tiny,length(pks)),2.5)),
             pks,pch=pltshp, bg=plotcol[j], cex=1.25)
    }
  }
}

legend(x="topleft", legend=c("RL","GB"), pch=c(21,24), 
       pt.bg="gray", cex=0.75, pt.cex=1.25)
legend(x="bottomleft",legend=species, pch=21, 
       pt.bg=plotcol[1:length(species)], cex=0.75, pt.cex=1.25)
```

#Graph to visualize AIC for Null and Alt for Year, Plot and Species
## Visual plot, not a great plot as its not by plot/species, but one can see AIC of Alt model is lower than null

#Second graph plots AIC null vs. AIC alt of all combos, and the 1:1 line. This should show most if not all AIC alt are less than AIC null. Could color code / change shapes depending on years / species
```{r}
#Make a plot showing...
aics_yrpltsp %>% pivot_longer(cols=c("AICnull","AICalt")) %>% mutate(value =as.numeric(value)) %>% ggplot(aes(x=year, y=value, fill=name,color=species,group=species)) +
    geom_bar(stat='identity', position='dodge') + theme_minimal(base_size = 14)+
  labs(fill='AIC value',y='AIC value',x="Year")

#Make a plot showing AIC null vs. AICalt
plot(aics_yrpltsp$AICnull, aics_yrpltsp$AICalt, pch=21, bg="grey",
     xlab="AIC Null", ylab="AIC Curve")
title("Compare Null to Curve")
abline(0,1)
```


#Explore Outliers
The following graphs are intended to detect outliers and model fitting issues

##Graph of plot / Species / Year-specific model fits
In this graph, the predicted peak flowering day of the year (DOY) of each species is graphed by year/plot id. X-axis is the years, Y-axis is DOY, and colors indicate all the species. This allows us to identify species / plot year fits that are anomolous (indicating potential model fitting issues) 
```{r}
plot_pars_years <- read_csv("data/PerPlotCurves.csv") 

# str_sub subsets first two characters pf the site ID, so here it would be either RL or GB.
# aes - aesthetics -we give x and y variables , and color 
# facets add third dimension to the figure here.
# %>%  is the pipe operator - which is equivalent to saying make all the field names available to the operation on right, it helps one to write simplified code where we don't need to use $ to access the fields in a dataframe

plot_pars_years %>%  mutate(site= str_sub(plot_pars_years$plot,1,2)) %>%
  ggplot() +
  geom_point(aes(as.factor(year),peak,color=species)) + 
  facet_grid(.~plot) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90,hjust = 5)) +
     geom_hline(aes(yintercept = 152))  +
  geom_text(mapping = aes(label = 'June 1st',y=152, x = 0.1), angle = 90,alpha=.4, hjust = 0) +
  geom_hline(aes(yintercept = 183))  +
  geom_text(mapping = aes(label = 'July 1st',y=183, x = 0.1), angle = 90, alpha=.4,hjust = 0) +
   geom_hline(aes(yintercept = 213))  +
  geom_text(mapping = aes(label = 'Aug 1st',y=213, x = 0.1), angle = 90, alpha=.4,hjust = 0) +
    geom_hline(aes(yintercept = 244))  +
  geom_text(mapping = aes(label = 'Sep 1st',y=244, x = 0.1), angle = 90,alpha=.4, hjust = 0) +
 # facet_grid(.~year) +
  labs(x=" Year", y= "Predicted Peak flowering (DOY)", color="Species")

# Its a comprehensive plot so might be better to view the generated image separately.
ggsave("figs/sanity_check.png",width = 70, height = 20, units = "cm")

``` 

#Alternative graph of model fits
Alternative plot to visualize outliers,  Site level (RL or GB) is on alternative y-axis, X-axis DOY, and Y-axis the focal species. We see the outliers are closer to start of the y-axis scale or towards the end. Columns here are years. SDD gradient can be ignored ( no variation)
```{r}
# note here aes  (aesthetics included 'fill' as well)
plot_pars_years %>% 
  mutate(site= str_sub(plot_pars_years$plot,1,2)) %>% 
  ggplot( aes(x = peak, y = species,fill=SDD)) +
   geom_jitter() +
    facet_grid(site~year) +
    scale_fill_viridis_c(name = "Snow Melt Date") +
    labs(title = 'Peaks across all the years',
       subtitle = 'Modelled peak flowering across 10 focal species', 
         x = "Day of the year") +
  theme_minimal(base_size = 14) + theme(axis.title.y = element_blank())

ggsave("figs/sanity_check_revised.png", width = 35, height = 20, units = "cm")
```

#Observation Effort
This code calculates the total number of observations for each plot / year / species combination.
```{r}

#This coce shows the total number of observations by plot / year / species
PhenoSite %>% 
  group_by(Transect,Year,Site_Code,Species)%>% 
  tally()

#This code does the same thing, but transforms the tibble into a dataframe
tallyObservations_typs <- PhenoSite %>% 
  group_by(Transect,Year,Site_Code,Species)%>% 
  tally() %>% as.data.frame()
```

#Graph showing species/ year / transect observer effort
Notice a perfect cluster of # of observationsin 2015 Reflections lake. Also, what we see is the low number of observations for some of the plot / species combinations

```{r}
# In dply lingo everything is a tibble , we are transforming it to a dataframe for compatibility
tallyObservations_typs %>% 
  ggplot() +
  geom_jitter(aes(as.factor(Year), n,color=Species))+
  facet_grid(.~Transect) +
  labs(y="Number of observations",x="Year") +
  theme_minimal(base_size = 14) 

```

#Calculate outlier 'yes' observations
This code calculates all pairwise differences in dates between all yes observations in each trail / year / plot / species combo, and identify any observation where the minimum value is > 7

For example, imagine a specific query : Species =='ANOC' & Year=='2013' & Site_Code =='RL9'

1.*group_by* orders the rows by given set of fields (in order)
2. *lag* gets the previous row

Returns

# Groups:   Transect, Year, Site_Code, Species [1]
  Transect          Year Site_Code Species Date      Observer                           QA.QC
  <chr>            <int> <chr>     <chr>   <chr>     <chr>                              <int>
1 Reflection Lakes  2013 RL9       ANOC    7/21/2013 Dave Purdon                            0
2 Reflection Lakes  2013 RL9       ANOC    7/23/2013 Anna Wilson; Cherry Chen               1
3 Reflection Lakes  2013 RL9       ANOC    7/23/2013 Weedy McCauley                         0
4 Reflection Lakes  2013 RL9       ANOC    7/24/2013 Brooke Upton                           0
5 Reflection Lakes  2013 RL9       ANOC    7/24/2013 Rita Moore; Dan Paquette               0
6 Reflection Lakes  2013 RL9       ANOC    7/25/2013 Carol Miltimore                        0
7 Reflection Lakes  2013 RL9       ANOC    7/26/2013 Carol Clingan                          0
8 Reflection Lakes  2013 RL9       ANOC    8/4/2013  Johndavid Hascup; Kristalyn Hascup     0
9 Reflection Lakes  2013 RL9       ANOC    8/10/2013 Bonnie Scott                           0

Row # 7 and 8 is what we use to find the difference which is greater than 7 

#TODO I suggest making pdif a variable we can play around with
#Question - does this only look at the difference between each observation and a previous flowering observation, or also the difference between previous and next? In the latter case, what happens to the first observation of flowering (no previous observation?)

```{r}
PhenoSite %>% 
  filter (Flower == 1) %>% 
  mutate (doy = lubridate::yday(as.Date(Date, "%m/%d/%Y"))  ) %>%
  group_by(Transect,Year,Site_Code,Species) %>%
  mutate(pdiff = doy -lag(doy, default = doy[1])) %>%
  filter (pdiff > 7) %>% 
  ggplot( aes(x = Site_Code, y = Species,fill=pdiff)) +
   geom_jitter() +
    facet_grid(Transect~Year)    +
  scale_fill_viridis_c(name = "Difference in DOY") +
    labs(title = 'Pairwise differences',
       subtitle = 'All species', 
         x = "Site Code") +
  theme_minimal(base_size = 14) + theme(axis.title.y = element_blank(),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

# Pairwise differences but filter it only by focal species
#I got rid of this snippet of code, as I changed code above to only look at focal species

# Plot pairwise for RL only, looks like 128 observations for RL (no jitter)

```{r}
PhenoSite %>% 
  filter (Transect== 'Reflection Lakes' & Flower == 1 & Species %in% species) %>%
  mutate (doy = lubridate::yday(as.Date(Date, "%m/%d/%Y"))  ) %>%
  group_by(Transect,Year,Site_Code,Species) %>% 
  mutate(pdiff = doy - lag(doy, default = doy[1])) %>%
  filter (pdiff > 7) %>% ggplot( aes(x = Site_Code, y = Species,fill=pdiff)) +
   geom_point() +
    facet_grid(Transect~Year)    +
  scale_fill_viridis_c(name = "Difference in DOY") +
    labs(title = 'Pairwise differences',
       subtitle = 'All species', 
         x = "Site Code") +
  theme_minimal(base_size = 14) + theme(axis.title.y = element_blank(),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

#Plot pairwise for GB only
```{r}
PhenoSite %>% 
  filter (Transect== 'Glacier Basin' & Flower == 1 & Species %in% species) %>%
  mutate (doy = lubridate::yday(as.Date(Date, "%m/%d/%Y"))  ) %>%
  group_by(Transect,Year,Site_Code,Species) %>% mutate(pdiff = doy -     lag(doy, default = doy[1])) %>%
  filter (pdiff > 7) %>% 
  ggplot( aes(x = Site_Code, y = Species,fill=pdiff)) +
   geom_jitter() +
    facet_grid(Transect~Year)    +
  scale_fill_viridis_c(name = "Difference in DOY") +
    labs(title = 'Pairwise differences',
       subtitle = 'All species (jittered)', 
         x = "Site Code") +
  theme_minimal(base_size = 14) + theme(axis.title.y = element_blank(),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
