---
title: "MW_WTASentiment.Rmd"
author: "Janneke Hille Ris Lambers, Aji John, Meera Sethi, Elli Theobald"
date: "12/22/2020"
output: html_document
---

**IN PROGRESS: exploratory in nature**
# This notebook expands on **MW_WTAanalysis.Rmd**


# Setup R markdown, load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(randomForest)
```

## Visualize the sentiments
```{r }
WTA_MW_dat <- read.csv("output/WTA_MW_dat.csv", header=TRUE)

WTA_MW_dat %>% 
  filter (trail=='GB') %>%
  ggplot() + geom_point(aes(year,total_s,color=month)) + geom_jitter(aes(year,total_s,color=month)) + theme_minimal(base_size = 18) +
  labs(x="Year", y="Overall score", color="Month")
```

## By year, and month what is the spread.
```{r, echo=FALSE}
WTA_MW_dat %>% 
  filter (trail=='GB') %>%
  ggplot() + geom_point(aes(year,month,color=as.factor(total_s))) + geom_jitter(aes(year,month,color=as.factor(total_s))) + theme_minimal(base_size = 18) +
 # scale_y_discrete(limits=c('January','May','June','July','August','September','October')) +
  labs(x="Year", y="Overall score", color="Score")
```

## By year, and month what is the spread - histogram
```{r, echo=FALSE}
WTA_MW_dat %>% 
  filter (trail=='GB') %>%
  ggplot() + geom_bar(aes(month,fill=as.factor(round(total_s,digits = 1)))) + 
  theme_minimal(base_size = 18) +
  facet_grid(.~year)+
  #coord_flip() +
  scale_x_discrete(limits=c('May','June','July','August','September','October')) +
  labs(x="Month", y="Count", fill="Score") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
## Overlap the sentiment score with richness

```{r, echo=FALSE}

Factor <- 1
WTA_MW_dat %>% 
  filter (trail=='GB') %>%
  ggplot() + 
  geom_bar(aes(month,fill=as.factor(round(total_s,digits = 1)))) +
  geom_smooth(aes(month,rich), method="loess", col="red") +
  theme_minimal(base_size = 18) +
  scale_y_continuous(name="Score", sec.axis=sec_axis(~./Factor, name="Richness")) +
  facet_grid(.~year)+
  #coord_flip() +
  scale_x_discrete(limits=c('May','June','July','August','September','October')) +
  labs(x="Month", y="Count", fill="Score") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

##
```{r, echo=FALSE}

Factor <- 1
WTA_MW_dat %>% 
  filter (trail=='GB') %>%
  ggplot() + 
  geom_bar(aes(DOY,fill=as.factor(round(total_s,digits = 1)))) +
  geom_smooth(aes(DOY,rich/10), col="red") +
  theme_minimal(base_size = 18) +
  facet_grid(.~year)+
  scale_y_continuous(name="Count", sec.axis=sec_axis(~./10, name="Richness")) +
  #coord_flip() +
  #scale_x_discrete(limits=c('May','June','July','August','September','October')) +
  labs(x="DOY", y="Count", fill="Score") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## See if the sentiment overlaps with species richness 
```{r, echo=FALSE}

scaleFactor = 2
WTA_MW_dat %>% filter (trail=='GB') %>%
  ggplot() + geom_point(aes(DOY,total_s)) + 
  geom_smooth(aes(DOY,total_s)) +
  geom_smooth(aes(DOY,y=rich/scaleFactor), method="loess", col="red") +
  scale_y_continuous(name="Score", sec.axis=sec_axis(~./scaleFactor, name="Richness")) +
  theme_minimal(base_size = 18) +
  facet_grid(.~year)+
  labs(x="DOY", y="Score",title = "GB - Richness vs Score")+
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
## Importance of individual sentiment scores on final sentiment
```{r }

WTA_MW_dat_notnull <- WTA_MW_dat[,c('bug_s','wild_s',
                                    'crowd_s','trail_s',
                                    'weather_s','view_s',
                                    'wildflower_s','total_s')]

bag.sentiments<-randomForest(total_s ~ . ,data=WTA_MW_dat_notnull,importance=TRUE,keep.inbag = TRUE,ntree=500,na.action = na.omit)
# View the forest results.
print(bag.sentiments) 
# Importance of each predictor.
print(importance(bag.sentiments,type = 2)) 
```

Trail sentiment contributes the most to the final score
```{r }
importrf <- importance(bag.sentiments,type = 2)
importrf <-importrf %>% as.data.frame()
importrf$variables <- c("bug_s",  "wild_s" ,  "crowd_s" ,  "trail_s" ,  "weather_s"  ,  "view_s",  "wildflower_s"  )
importrf<-importrf%>% arrange(desc(IncNodePurity) )
importrf%>% ggplot() + geom_col(aes(x=reorder(variables, -IncNodePurity),IncNodePurity)) +
  theme_minimal(base_size = 18) +
  labs(x="Variables","Importance") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Is it - Trail sentiment is correlated with Wildflower sentiment 

```{r }
rel_trail_wildflower <- lm(data = WTA_MW_dat_notnull,trail_s~wildflower_s)

summary(rel_trail_wildflower)
```

Not really.


## check if trail sentiment was sufficient to explain the total sentiment

```{r }
rel_total_trail <- lm(data = WTA_MW_dat_notnull,total_s~trail_s)
summary(rel_total_trail)
```
Singificant, and explains 15% of the variation.

## Lets check the richness and total sentiment

```{r }

WTA_MW_dat_richness <- WTA_MW_dat[,c('bug_s','wild_s',
                                    'crowd_s','trail_s',
                                    'weather_s','view_s',
                                    'wildflower_s','total_s','total_s2','rich')]
rel_total_richness <- lm(data = WTA_MW_dat_richness,total_s~rich)

summary(rel_total_richness)
```
Not really correlated.


## Visitation and sentiment related ?
```{r }

MR_Visitation <- read_csv('./data/Visitation by Month.csv')
# Reshaping the dataframe -
MR_Visitation_reshaped <- MR_Visitation %>%
 pivot_longer(-Year, names_to = "Month", values_to = "Visitation")
head(MR_Visitation_reshaped)
```

```{r }
MR_Visitation_reshaped %>% 
  filter(Year %in% c(2015:2019)) %>% 
  ggplot() + geom_point(aes(Month,Visitation, color=as.factor(Year))) +
   scale_x_discrete(limits=c('JAN','FEB','MAR','APR','MAY',
                             'JUN','JUL','AUG','SEP','OCT','NOV','DEC')) +
  theme_minimal(base_size = 18)+
  labs(color='Year') +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Highest visitations in 2019, especially in peak months - July and August. Seems to link to trail sentiments. 

