---
title: "Wildflower Season from MeadoWatch Data"
author: "Janneke Hille Ris Lambers, Aji John, Meera Sethi, Elli Theobald"
date: "10/01/2020"
output: html_document
---

#Setup for R script and R markdown
1. Load all libaries
2. Specify string behavior
3. Load packages (should we add all packages here?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = FALSE)
library(tidyverse)
library(leaflet)
library(lubridate)
library(boot)
library(readr) 
```


#Data munging / merging
1. Read in raw data
2. Remove cases with NA observations in any phenophase
3. Merge plot-specific snowmelt data and individual observations
4. Remove unneeded columns
5. Calculate and add Julian day (DOY) to data
6. Reorganize data columns and rename to PhenoSite

```{r}
# Read in phenology data, station data
PhenoDatall <- read.csv("data/MW_PhenoDat_2013_2019.csv", header=TRUE)
StationDat <- read.csv("data/MW_SiteDat_2013_2019.csv", header=TRUE) 

# Remove rows where any phenophases were coded as NA
PhenoDat <- PhenoDatall[complete.cases(PhenoDatall[,c(12, 14, 16, 18)]),] 

# Merge by the rows in both the files (Year, Site_Code)
MergePhenoStation <- merge(StationDat,PhenoDat, by=c("Year","Site_Code"))

# Remove unneeded columns
#TODO - might be a better way to do this in case column names change
PhenoSite_0 <- MergePhenoStation[,c(1:6, 9, 12, 15:16, 18:20, 22, 24, 26, 28)]

# Calculate Julian Days of observations; DSS=days since snow; add to PhenoSite
PhenoSite_0 <- PhenoSite_0[order(PhenoSite_0$Year),] #order by year
yrs <- unique(PhenoSite_0$Year);
DOY <- c()

# Convert the 'Date' - observed date to Julian (DOY)
for(i in 1:length(yrs)){
  tmpdates <- PhenoSite_0$Date[PhenoSite_0$Year==yrs[i]]
  Jan1Jul <- as.Date(paste(yrs[i],"-01-01", sep=""))
  ObsJulDayYr <- julian(as.Date(as.character(tmpdates),"%m/%d/%Y"),
                        origin=Jan1Jul)
  DOY <- c(DOY,ObsJulDayYr)
}

# Now add day of year (DOY) to Phenosite Data
PhenoSite_0 <- cbind(PhenoSite_0, DOY)

# Finally, reorganize data and rename as PhenoSite
PhenoSite <- PhenoSite_0[,c(1,3,2,4,8,18,7,10,11,14)] #reorganize data

#Examine the data
head(PhenoSite)

```


#Define functions - maximum likelihood models & phenology curves (for plotting)
There are four functions we need, 3 for model fitting and one for drawing curves. They are described here.

1. nullfit - binomial likelihood model that estimates a constant probability of flowering (one parameter) and the negative log likelihood, given yes / no observations. We know this model isn't accurate, but it is useful in providing
a null model for curvefit_perplot (allowing one to compare AIC's or conduct a likelihood test as to whether flowering varies with DOY in a unimodal way).

2. curvefit_perplot - binomial likelihood model that estimates the probability of flowering as a function of DOY; as described by three parameters - range, maximum, and peak. The unimodal curve is fit with a logit transformation (similar to Theobald et al 2017, Sethi et al 2020; albeit with parameters that do NOT vary with climate). Also requires yes / no observations. This function is fit to year-plot-species specific observations

3. curvefit_allplot - binomial likelihood model that estimates the probability of flowering for a species across all plots, years and transects as a function of several parameters. Similar to the second model above, there is a unimodal curve describing flowering, fit with a logit transformation. Peak flowering at a plot is predicted by snowmelt date (intercept, slope parameter estimated). Range and maximum parameters vary by year and plot. This function is fit to all data from a species.

4. predflower - input a DOY (or vector of DOY's) and 3 parameters describing peak, range and duration of flowering; and returns the probability of flowering. This is primarily used for plotting.

```{r}
#Null model - assumes the probability of flowering is constant
nullfit <- function (param){ 
	meanp  <- param[1]
	pred   <- rep(meanp, times=length(phenophase))
	llik     <- dbinom(phenophase,1,pred, log=TRUE) #this is the likelihood
	return(-sum(llik)) #this is the negative log likelihood, which is minimized
}

#Assumes flowering varies with DOY as a function of 3 parameters
#Should be fit per species, plot and year
curvefit_perplot <- function (param){ #curve fitting function for mle
	 peakp  <- param[1]
	 rangep <- param[2]
	 maxp   <- param[3]
	 pred <- inv.logit(rangep * (days - peakp)^2 + maxp)
	 llik     <- dbinom(phenophase,1,pred, log=TRUE)
	 return(-sum(llik))
}

#Fits model per species 
#peak flowering varies with snowmelt, and plot/trail specific range and max
#Should be fit per species (but all data for each speices)
curvefit_allplot <- function (param){ #curve fitting function for mle
  intpeak  <- param[1]
  slopepeak <- param[2]
  rangep_all <- param[3:(2+ntrlyr)] #different range per trail, year
  maxp_all   <- param[(3+ntrlyr):(2+2*ntrlyr)] #different max per trail, year
  peakp <- intpeak + slopepeak*SDD
  rangep <- rangep_all[trlyr]
  maxp <- maxp_all[trlyr]
  pred <- inv.logit(rangep * (days - peakp)^2 + maxp)
  llik     <- dbinom(phenophase,1,pred, log=TRUE)
  return(-sum(llik))
}

#Predicts flowering phenology as a function of DOY, param (peak, range, max)
predflower <- function (xx, param){ 
  days    <- xx
	peakp  <- param[1]
	rangep <- param[2]
	maxp   <- param[3]
	pred <- inv.logit(rangep * (days - peakp)^2 + maxp)
	return(pred)
}
```


#Define wildflower community for model fitting
This chunk of code defines the wildflower community and subsets the data to only include these species. Specifically:

1. Specify which flowering species to include. For reference:
   A. Reflection Lakes species include: c("ANOC", "CAPA", "ERMO", "ERPE", "LIGR", "LUAR", "MIAL", "PEBR", "POBI", "VASI")
   B. Glacier Basin species include: c("ANAR", "ARLA", "ASLE", "CAMI", "ERGR", "LUAR", "MEPA", "PEBR", "POBI", "VASI")

2. Create PhenoSite_Focal - data set with just species of interest

3. From this, *nspp* (number of species) and *nyears* (number of years) are defined, which are used in for loops

```{r}
# Define the wildflower community: which species to include
species <- c("ANOC", "ASLE", "CAMI", "CAPA", "ERGR", "ERMO",
             "ERPE", "LUAR", "PEBR", "VASI")

# Create PhenoSite_Focal - data set with just species of interest
specieskeep <- PhenoSite$Species %in% species
PhenoSite_Focal <- PhenoSite[specieskeep,]

# Determine how many years of data, how many species, etc
yrs <- unique(PhenoSite_Focal$Year)
nyrs <- length(yrs)
nspp <- length(species)

```


#Phenological models fit to year-plot-species specific data
This chunk of code uses several for loops to fit phenological curves to year-plot-species specific data (using function *curvefit_perplot*). The advantage of these curves is that they are optimized to individual plots. The disadvantage is that these curves only exist if the species was present in the plot (even if it was nearby). More broadly, these fits can be used to check output from likelihood models fit to all data from a species.

##Code does the following:
1. Pulls out data for each species / plot / year combo. Data in innermost loop is PhenoSite_YearPlotSpecies. Note - fitting only occurs if sufficient data, as defined by these two parameters (you can modify them):
   + *totobs* (# observations for fitting, set at 10)
   + *totyesobs* (# observations of flowering, set at 5)

2. The functions nullfit and curvefit_perplot are fit to year-plot-species specific data.

3. AIC's are calculated for both the null model and curve fitting model, which can be used to determine whether time (DOY) helps explain the probability of flowering. This information is currently not saved.

4. Plots individual curves and raw data in separate graphics windows (if *plottrue* is set to "T")

5. Saves the peak, range and maximum parameters to object pars_yrpltsp

6. Turn pars_yrpltsp into a data frame (this happens after the for loop)

```{r}
#define object in which to save parameters
pars_yrpltsp <- c() 

#set plottrue to 'T' if plots desired; 'F' if not
plottrue <- "F"

#set data filters - how many total obs, how many flowering obs
#to proceed with model fitting?
totobs <- 20 #need at least 20 observations (visits) 
totyesobs <- 5 #flowering needs to be observed at least 5 times 

#For loops to fit curves per year/plot/species, save parameters, plot 
for(i in 1:nyrs){ # First for loop: runs through years

  #extract data for that year
  PhenoSite_Year <- PhenoSite_Focal[PhenoSite_Focal$Year==yrs[i],]
  
  #pull out unique plots for year in question
  plots <- unique(PhenoSite_Year$Site_Code)
  
  #Nested for loop for each plot
  for(j in 1:length(plots)){ # Second for loop: each plot
    PhenoSite_YearPlot <- PhenoSite_Year[PhenoSite_Year$Site_Code==plots[j],]

    #Identify species in plot
    spinplt <- unique(PhenoSite_YearPlot$Species)
    if(length(spinplt)==0){next} #break if focal spp not in plot
    
    #Nested for loop for each species
    for(k in 1:length(spinplt)){ #Third for loop: each species in the pot
      
      #Set up graphics windows for multiple plots, if plottrue set to "T"
      if(plottrue=="T"){
        if(i==1&j==1&k==1){ #set up graphics window, if plotting
          X11(width=8,height=8)
          par(mfrow=c(4,4), tck=-0.02, omi=c(0,0,0,0), 
              mai=c(0.4,0.4,0.4,0.2), mgp=c(1.25,0.5,0))
          newplot <- 0
        }

        if(newplot==16){
          X11(width=8,height=8)
          par(mfrow=c(4,4), tck=-0.02, omi=c(0,0,0,0),
              mai=c(0.4,0.4,0.4,0.2), mgp=c(1.25,0.5,0))
          newplot <- 0
        }
      }
      
      #Extract data for the species in question      
      PhenoSite_YearPlotSpecies <- 
        PhenoSite_YearPlot[PhenoSite_YearPlot$Species==spinplt[k],]
      
      #define parameters for curvefitting
      days <- PhenoSite_YearPlotSpecies$DOY #explanatory variable: DOY 
      phenophase <- PhenoSite_YearPlotSpecies$Flower #yes / no flowering

      #Do not fit model to data if insufficient data
      if(length(days)<=totobs){next}
      if(sum(na.omit(phenophase))<=totyesobs){next}
      
      #remove days when no observations were made; NA in phenophase
      #TODO is this redundant with cases line? check with earlier code
	    days <- days[is.na(phenophase)==FALSE]
	    phenophase <- phenophase[is.na(phenophase)==FALSE]
	    
	    #add three weeks of zeroes before earliest SDD in those plots
	    SDDplt <- min(PhenoSite_YearPlotSpecies$SDD)
	    days <- c(SDDplt-21, SDDplt-14, SDDplt-7,days)
	    phenophase <- c(0,0,0,phenophase)

      #now fit null model
      model0 <- optimize(nullfit, c(0.000001,0.999999)) #fit null model
	
      #now fit alternative model - curve
      param <- c(mean(days[phenophase[]==1]), -0.001, 0) # initial parameters
      model1 <- optim(param, curvefit_perplot, control = list(maxit = 50000))
      if(model1$convergence==1){
        print(paste(spinplt[k],"no convergence", sep="-"))}
    	
      #plot curve, data only if plottrue is set to "T"
      if(plottrue == "T"){
        plot(days,phenophase, ylab="flower",pch=21, bg="grey")
        xx <- seq(min(days),max(days))
        yy <- predflower(xx,model1$par)
        lines(xx,yy)
        title(paste(nyears[i],plots[j],spinplt[k],sep=" "))
        newplot <- newplot + 1
      }
      
      #TODO 
      #write the model tests / AIC to data frame
      AICnull <- round(2*(model0$objective+1),1)
      AICalt <- round(2*(model1$value + 3),1)
      pcurve <- signif(pchisq(model1$value-model0$objective,2),3)

      #save parameters to spyrpltpars
      tmp_pars <- c(yrs[i], as.character(plots[j]), SDDplt, 
                    as.character(spinplt[k]), model1$par[1:3])
      pars_yrpltsp <- rbind(pars_yrpltsp, tmp_pars)
    }
  }
}
  
# turn spyrpltpars into a data frame
dimnames(pars_yrpltsp) <- list(c(), c("year", "plot", "SDD", "species",
                                     "peak", "duration", "max"))
pars_yrpltsp <- data.frame(pars_yrpltsp)

#change storage type to numeric - all except plot (since a few plots have a, b)
pars_yrpltsp$year <- as.numeric(pars_yrpltsp$year)
pars_yrpltsp$species <- as.factor(pars_yrpltsp$species)
pars_yrpltsp$SDD <- as.numeric(pars_yrpltsp$SDD)
pars_yrpltsp$peak <- as.numeric(pars_yrpltsp$peak)
pars_yrpltsp$duration <- as.numeric(pars_yrpltsp$duration)
pars_yrpltsp$max <- as.numeric(pars_yrpltsp$max)

head(pars_yrpltsp)
```


#Graph to visualize peak flowering as observed by MW volunteers each year
This code takes estimates of peak flowering from year-plot-species specific model fits (in pars_yrpltsp) and plots those estimates by year (year on the x-axis, DOY on the y axis) and by trail.

```{r}
##Plot peak flowering estimates of all species, trails, plots on one graph
par(mfrow=c(1,1),omi=c(0,0,0,0), mai=c(0.5,0.4,0.4,0.2), 
    tck=-0.01, mgp=c(1.25,0.25,0), xpd=TRUE)

#set plotting colors - 16 total to accommodate all species if needed
plotcol <- c("yellowgreen","magenta","orange","purple","yellow","springgreen",
             "pink","purple","navyblue","azure4","yellow4","orchid",
             "turquoise","salmon","maroon","black")

#create plot to add points to
#earlypk <- min(spyrpltpars$peak); latepk <- max(spyrpltpars$peak)
earlypk <- 135; latepk <- 255

plot(2016,175, xlim=c(2012, 2020), ylim=c(earlypk,latepk),type="n",
     xaxp=c(2012,2020,8), yaxt="n", xlab="Year",ylab="Flowering")
text(2011.5,srt=90, c(135, 165, 195, 225, 255),-0.1, 
     labels=c("May", "Jun", "Jul", "Aug", "Sept"))
  
for(trail in 1:2){
  if(trail==1){
    pars_yrpltsp2 <- pars_yrpltsp[substr(pars_yrpltsp$plot,1,2)=="RL",]}
  if(trail==2){
    pars_yrpltsp2 <- pars_yrpltsp[substr(pars_yrpltsp$plot,1,2)=="GB",]}
  
  #how many years? differs per trail
  yrs2 <- unique(pars_yrpltsp2$year)

  #extract data per year, plot
  for(i in 1:length(yrs2)){
    paryear <- pars_yrpltsp2[pars_yrpltsp2$year==yrs2[i],]
 
    #now pull out all data for a species
    for(j in 1:length(species)){
      paryearsp <- paryear[paryear$species==species[j],]
      if(dim(paryearsp)[1]==0){next}
      pks <- paryearsp$peak
      tiny <- 0; if(trail==2){tiny <- 0.333}
      if(trail==1){pltshp <-21}; if(trail==2){pltshp <- 24}
      points((rep(yrs2[i],length(pks))+jitter(rep(tiny,length(pks)),2.5)),
             pks,pch=pltshp, bg=plotcol[j], cex=1.25)
    }
  }
}

legend(x="topleft", legend=c("RL","GB"), pch=c(21,24), 
       pt.bg="gray", cex=0.75, pt.cex=1.25)
legend(x="bottomleft",legend=species, pch=21, 
       pt.bg=plotcol[1:length(species)], cex=0.75, pt.cex=1.25)
```


#Fit phenology model for all species data
This code fits a predictive phenology model to all observations of flowering for a particular species (function *curvefit_allplot*), for the same flowering community as defined above. So, this code chunk expects that the data frame *PhenoSite_Focal*, the vector *species*, and the objects *nspp*, *totobs* and *totyesobs* have been defined. Note that this takes a bit of time to run (5ish minutes). The code does the following:

1. Subset data for a particular species.

2. For each species, use a for loop to assemble year-plot specific data for the species in question. This requires extracing year-plot specific data, assessing whether there is sufficient data (so only plots with *totobs* observations and *totyesobs* flowering observations) are included, and adding three no flowering observations before snowmelt (as was done for model fitting for individual year-plot-species data), and creating variables *days* (vector of DOY for all plots included), *phenophase* (vector of 1's and 0's observed), *SDD* (snow disappearance date at each plot for the observation in question) and *trlyr* which trail / year combo the observation came from (this is for fitting range, max parameters).

3. Fit the maximum likelihood model (defined by *curvefit_allplot* above) to the data. This requires setting initial parameters first - which was done by fitting the relationship between peak flowering and SDD from year-plot-species specific data (in *pars_sppltsp*), and taking the average range and duration values for those same model fits.

4. Saves the parameters to *pars_spp*

5. After the for loop, turns *pars_spp* into a data frame

```{r}
#define where to save output
pars_spp <- c() #save parameters for per species fit, peak f based on SDD

##Now nested for loops to fit curves for all years, species, plots
for(i in 1:nspp){ #loop for each species, to assemble data
  #extract data for that year
  PhenoSite_Species <- PhenoSite_Focal[PhenoSite_Focal$Species==species[i],]
  
  #assemble data - need to add zeros before SDD, define trail year combos
  yrplt <- paste(PhenoSite_Species$Year, PhenoSite_Species$Site_Code, sep="")
  yrplts <- unique(yrplt)
  yeartrail <- unique(substr(yrplts,1,6))
  yeartrail <- yeartrail[order(yeartrail)]
  ntrlyr <- length(yeartrail)
  trlyr <- c()
  days <- c()
  phenophase <- c()
  SDD <- c()
  
  #Now go through each unique year - plot combination per species
  for(j in 1:length(yrplts)){
     PhenoSite_YearPlotSpecies <- PhenoSite_Species[yrplt[]==yrplts[j],]
     tmpdays <- PhenoSite_YearPlotSpecies$DOY #explanatory variable: DOY 
     tmpphenophase <- PhenoSite_YearPlotSpecies$Flower #yes / no flowers
     tmpSDD <- PhenoSite_YearPlotSpecies$SDD
     
     #Do not include if fewer totobs obs, totyesobs flowering obs
     if(length(tmpdays)<=totobs){next}
     if(sum(tmpphenophase)<=totyesobs){next}
     
     #remove days when no observations were made; NA in phenophase
     tmpdays <- tmpdays[is.na(tmpphenophase)==FALSE]
     tmpSDD <- tmpSDD[is.na(tmpphenophase)==FALSE]
     tmpphenophase <- tmpphenophase[is.na(tmpphenophase)==FALSE]
     
     #Add observations of no flowering 3 weeks prior to snowmelt
     SDDplt <- min(tmpSDD)
     tmpdays <- c(SDDplt-21, SDDplt-14, SDDplt-7,tmpdays)
     tmpphenophase <- c(0,0,0,tmpphenophase)
     tmpSDD <- c(tmpSDD[1:3],tmpSDD)
    
     #determine which trail year combo
     whichtrlyr <- substr(paste(PhenoSite_YearPlotSpecies$Year[1],
                                PhenoSite_YearPlotSpecies$Site_Code[1], 
                                sep=""),1,6)
     tmptrlyr <- rep(which(whichtrlyr==yeartrail[]),times=length(tmpdays))
     #now append to days, phenophase,trlyear
     days <- c(days, tmpdays)
     phenophase <- c(phenophase, tmpphenophase)
     SDD <- c(SDD, tmpSDD)
     trlyr <- c(trlyr, tmptrlyr)
  }
  #now fit curvefit_all plot model - optim as function of SDD
  
  #first use individual year-plot-species to come up with starting points
  #slope & intercept for SDD vs optim
  optimcoef <- unlist(coef(lm(pars_yrpltsp$peak~pars_yrpltsp$SDD)))
  spdur <- median(pars_yrpltsp$duration[pars_yrpltsp$species==species[i]])
  spmx <- median(pars_yrpltsp$max[pars_yrpltsp$species==species[i]])

  #specify initial parameters (note, optim can be fussy)
  param <- c(optimcoef, rep(spdur, times=ntrlyr), rep(spmx, times=ntrlyr))
  modelsnowmod <- optim(param, curvefit_allplot, control = list(maxit = 50000))

  #Determine which years / trails (for plotting, saving)
  trails <- "both"
  if(length(unique(substr(yeartrail,5,6)))==1){
    if(unique(substr(yeartrail,5,6))=="RL"){trails <- "Reflection Lakes"}
    if(unique(substr(yeartrail,5,6))=="GB"){trails <- "Glacier Basin"}
  }
 
  #Save parameters to pars_spp
  if(trails=="Reflection Lakes"){
    tmppars <- c(modelsnowmod$par[1:2], rep(NA, times=(ntrlyr-2)),
                 modelsnowmod$par[3:(2+ntrlyr)],rep(NA, times=(ntrlyr-2)),
                 modelsnowmod$par[(3+ntrlyr):(2+ntrlyr+ntrlyr)])
  }
  if(trails=="Glacier Basin"){
    tmppars <- c(modelsnowmod$par[1:2], modelsnowmod$par[3:(2+ntrlyr)], 
                 rep(NA, times=(ntrlyr+2)),
                 modelsnowmod$par[(3+ntrlyr):(2+ntrlyr+ntrlyr)],
                 rep(NA, times=(ntrlyr+2)))
  }
  if(trails=="both"){
    tmppars <- modelsnowmod$par
  }
  tmppars2 <- c(species[i],tmppars)
  pars_spp <- rbind(pars_spp,tmppars2)
}

#save original;
pars_spp2 <- pars_spp

#make pars_spp a data frame
dimnames(pars_spp) <- list(c(), c("species","optimint", "optimslope",
                                  "dur-GB15","dur-GB16","dur-GB17","dur-GB18",
                                  "dur-GB19","dur-RL13","dur-RL14","dur-RL15",
                                  "dur-RL16","dur-RL17","dur-RL18","dur-RL19",
                                  "max-GB15","max-GB16","max-GB17","max-GB18",
                                  "max-GB19","max-RL13","max-RL14","max-RL15",
                                  "max-RL16","max-RL17","max-RL18","max-RL19"))
pars_spp <- data.frame(pars_spp)

#change storage type to numeric - all except plot (since a few plots have a, b)
pars_spp$species <- as.factor(pars_spp$species)

for(i in 2:27){
  pars_spp[,i] <- as.numeric(pars_spp[,i])
}

head(pars_spp)

```


#Graph meadow level species richness by year / trail
This code graphs the probability of observing flowering at the meadow scale (i.e. when at one location along the trail) depending on the DOY, for the years / trails of MW. This function takes output from species specific models fit to all data, and uses the mean observed snowmelt date for that year / trail.

```{r}

```


#Predict species richness on the trails by DOY (IN PROGRESS)
```{r}
#For loop per year, and then trail
for(i in 1:nyrs){
  if(trail==1){
    StationDatTrail <- StationDat[StationDat$Transect=="Reflection Lakes",]
    parameters_snowmodtrail <- parameters_snowmod[is.na(parameters_snowmod$dur.RL13)==FALSE,]
    parameters_snowmodtrail <- cbind(parameters_snowmodtrail[,1:3],parameters_snowmodtrail[,substr(dimnames(parameters_snowmodtrail)[[2]],5,6)=="RL"])
    }
  if(trail==2){
    StationDatTrail <- StationDat[StationDat$Transect=="Glacier Basin",]
    parameters_snowmodtrail <- parameters_snowmod[is.na(parameters_snowmod$dur.GB15)==FALSE,]
    parameters_snowmodtrail <- cbind(parameters_snowmodtrail[,1:3],parameters_snowmodtrail[,substr(dimnames(parameters_snowmodtrail)[[2]],5,6)=="GB"])
  }
  
  #Now cycle through years
  predDOY <- seq(105,255)
  yrs <- unique(StationDatTrail$Year)
  
  for(j in 1:length(yrs)){
    SDDs <- StationDatTrail$SDD[StationDatTrail$Year==yrs[j]]
    firstSDD <- min(SDDs); lastSDD <- max(SDDs)

    #Set plotting parameters
    X11(width=6,height=8)
    par(mfrow=c(2,1), omi=c(0,0,0,0), mai=c(0.5,0.4,0.4,0.3), xpd=NA, tck=-0.02,mgp=c(1.25,0.5,0))
    
    #now extract parameters per species, plot
    parameters_smyear <- parameters_snowmodtrail[,c(1:3,(3+j),(3+j+length(yrs)))]
    spp <- parameters_smyear$species
    
    #first plot early season
    plot(160,0.5,type="n", xlim=c(min(predDOY),max(predDOY)), ylim=c(0,1),
         xaxp=c(150,270,5), xaxt="n", xlab="Time",ylab="Flowering")
    text(c(105,135,165,195,225,255),-0.1, labels=c("Apr","May","Jun","Jul","Aug","Sep"))
    if(trail==1){title(paste("Early meadows","RL",yrs[j],sep="-"))}
    if(trail==2){title(paste("Early meadows","GB",yrs[j],sep="-"))}
    
    #now plot early season
    for(k in 1:length(spp)){
      peakfirst <- parameters_smyear[k,2]+parameters_smyear[k,3]*firstSDD
      durf <- parameters_smyear[k,4]
      maxf <- parameters_smyear[k,5]
      
      #estimate flowering curve first meadows
      param <- c(peakfirst,durf,maxf)
      flfirst <- predphen(predDOY,param)

      #now plot
      lines(predDOY[flfirst[]>0.01],flfirst[flfirst[]>0.01],col=plotcol[k],lwd=2)
    }
    #add legend
    legend(x="topleft", legend=spp, cex=0.75, col=plotcol[1:length(spp)],lwd=2)

    #next plot late season
    plot(160,0.5,type="n", xlim=c(min(predDOY),max(predDOY)), ylim=c(0,1),
         xaxp=c(150,270,5), xaxt="n", xlab="Time",ylab="Flowering")
    text(c(105,135,165,195,225,255),-0.1, labels=c("Apr","May","Jun","Jul","Aug","Sep"))
    if(trail==1){title(paste("Late meadows","RL",yrs[j],sep="-"))}
    if(trail==2){title(paste("Late meadows","GB",yrs[j],sep="-"))}
    
    #now plot late season species
    for(k in 1:length(spp)){
      peaklast <- parameters_smyear[k,2]+parameters_smyear[k,3]*lastSDD
      durf <- parameters_smyear[k,4]
      maxf <- parameters_smyear[k,5]
      
      #estimate flowering curve last
      param <- c(peaklast,durf,maxf)
      fllast <- predphen(predDOY,param)
      
      #now plot
      lines(predDOY[fllast[]>0.01],fllast[fllast[]>0.01],col=plotcol[k],lwd=2)
    }
  }
}

```

